---
title: "R Notebook"
output: html_notebook
---

# Loading Data into R
```{r}
# Get your working directory in order with getwd() and setwd("")
d <- read.csv("Dataset_S1.txt") # Load data
```

# Exploring and Transforming Dataframes
```{r}
head(d, n=3) 
# Check data

nrow(d) 
# Check # rows

ncol(d) 
# Check # columns

dim(d) 
# Check dimensions (r c)

colnames(d) 
# Check column names for discrepancies; "X.GC" instead of "%GC"

colnames(d)[12] <- "Percent.GC" 
# Changed X.GC

d$depth 
# Access a single column of a dataframe with $

mean(d$depth) 
# The mean of one column in d

summary(d$depth) 
# Summary stats of one column in d

d[ , 1:2] 
# Subsetting columns

d[, c("start","end")] 
# We can also use column names to acheive the same end

d[1,c("start","end")] 
# Only first row of start and end columns

d[, "start", drop=FALSE]
# Disables subsetting as vector (retains dataframe format)
d$cent <- d$start >= 25800000 & d$end <= 29700000 
```

## Creating Dataframes From Scratch
```{r}
x <- sample(1:50,300,replace=TRUE) 
# Samples numbers 1-50, 300 times

y <- 3.2*x + rnorm(300,0,40) 
# Function

d_sim <- data.frame(y=y, x=x) 
# Represents vectors as a dataframe
```


# Logical vector for determining if window within a centromeric region
table(d$cent) # See how many rows were within the centromeric region
sum(d$cent) # Another way to see how many rows were within region
d$diversity <- d$Pi / (10*1000) # Rescaling data

## Exploring Data Through Slicing and Dicing
summary(d$total.SNPs)
# SNPs are heavily skewed right: Q3 = 12 but Max = 93! Let's investigate
d$total.SNPs >= 85 # Create logical vector
d[d$total.SNPs >= 85, ] # Subset according to logical vector
# Only rows that have a TRUE value (total.SNPs >= 85) will be kept.
d[d$Pi > 16 & d$Percent.GC > 80, ] # A more elaborate subsetting
d[d$Pi > 16 & d$Percent.GC > 80, c("start","end","depth","Pi")]
# One more time but asking for specific columns
d$Percent.GC[d$Pi > 16] 
# Asking for a specific column using a logical vector fromn ANOTHER COLUMN!!!
summary(d$depth[d$Percent.GC >= 80])
# Subsetting to summarize average depth in the >= 80 window; it is low compared to:
summary(d$depth[d$Percent.GC < 80])

summary(d$Pi[d$cent])
# Looking at Pi by windows falling in our above centromere window...
summary(d$Pi[!d$scent])
# ...and those that fall outside it.

d$Pi > 3
# which() takes a logical vector and returns the positions of all TRUE values!
which(d$Pi>3)
which(d$Pi>10)[1:4] # First four TRUE values
d[which.min(d$total.Bases),] # Returns row which has the minimum element
d[which.max(d$depth),] # Returns row with the maximum element
# So the centromere appears to have higher nucleotide diversity than other regions

# Sometimes subsetting expressions in brackets is redundant so we can use subset()
# subset() takes two arguments: dataframe and conditions to include a row:
subset(d,Pi > 16 & Percent.GC > 80)
# A third argument supplied to specify which columns to include:
subset(d, Pi > 16 & Percent.GC > 80,
       c(start,end,Pi,Percent.GC,depth))
# subset is special in that we don't have to "quote" column names



# Don't forget to git push this once done